### 2024-08-02 data-exercise/data-exercise.qmd
redid repos

---
title: "Data Exercise"
author: "Leonel Salazar, IGE436"
---

# I am not 100% done, i still need to add comments and analysis for the models. I will plan on doing this by the end of the 4th of july weekend.

```{r}

# Load required packages
library(ggplot2)  # For advanced plotting
library(dplyr)    # For data manipulation
library(tidyr)    # For data tidying
library(tidyverse)

```

# Set seed for reproducibility

```{r}

set.seed(789)

```

# Number of observations

```{r}

n <- 3000

```

# Generate variables

```{r}

Gender <- sample(c("Male", "Female"), n, replace = TRUE)
Age <- round(rnorm(n, mean = 45, sd = 10))
Income <- round(rnorm(n, mean = 60000, sd = 20000))
Education <- sample(c("High School", "Bachelor's", "Master's", "PhD"), n, replace = TRUE, prob = c(0.25, 0.4, 0.25, 0.1))
HomeOwner <- sample(c("Yes", "No"), n, replace = TRUE, prob = c(0.7, 0.3))
Retired <- sample(c("Yes", "No"), n, replace = TRUE, prob = c(0.15, 0.85))

```

# Combine into a data frame

```{r}

data <- data.frame(Gender, Age, Income, Education, HomeOwner, Retired)

```

# Summary and structure of the dataset

```{r}

summary(data)
str(data)

```

```{r}

# Convert 'HomeOwner' from character to factor with levels "No" and "Yes"
data$HomeOwner <- factor(data$HomeOwner, levels = c("No", "Yes"))

# Show structure of 'HomeOwner' after conversion
str(data$HomeOwner)

data$Retired <- factor(data$HomeOwner, levels = c("No", "Yes"))

```

```{r}

# Frequency table for categorical variables
table(data$Gender)
table(data$Education)
table(data$HomeOwner)
table(data$Retired)

```

# Created histogram and box plot to check for normality and outliers.
```{r}

# Histograms for Age and Income
par(mfrow=c(1,2))  # Arrange plots in one row, two columns
# Histogram for Age
hist(data$Age, 
     main = "Age Distribution", 
     xlab = "Age", 
     ylab = "Frequency",
     col = "skyblue",  # Histogram color
     border = "black"  # Border color
)

# Histogram for Income
hist(data$Income, 
     main = "Income Distribution", 
     xlab = "Income", 
     ylab = "Frequency",
     col = "lightgreen",  # Histogram color
     border = "black"     # Border color
)

# Box plot for Age
boxplot(data$Age,
        main = "Box Plot of Age",
        ylab = "Age",
        col = "skyblue",  # Box color
        border = "black", # Border color
        horizontal = FALSE  # Vertical box plot
)

# Box plot for Income
boxplot(data$Income,
        main = "Box Plot of Income",
        ylab = "Income",
        col = "lightgreen",  # Box color
        border = "black",    # Border color
        horizontal = FALSE   # Vertical box plot
)

```
# Barplots to explore visually the categorical variables.

```{r}

# Bar plots for categorical variables
par(mfrow=c(2,2))  # Arrange plots in a 2x2 grid
barplot(table(data$Gender), main="Gender Distribution")
barplot(table(data$Education), main="Education Level")
barplot(table(data$HomeOwner), main="Home Ownership")
barplot(table(data$Retired), main="Retirement Status")

```

```{r}

# Model 1: Income ~ Age
model1 <- lm(Age ~ Age, data = data)

# Summary of the model
summary(model1)

str(data)

```

```{r}

# Model 2: Income ~ Age + Gender
model2 <- lm(Income ~ Age + Gender, data = data)

# Summary of the model
summary(model2)

```

```{r}

# Model 3: HomeOwner ~ Age + Income + Education
model3 <- glm(HomeOwner ~ Age + Income + Education, data = data, family = binomial)

# Summary of the model
summary(model3)

```

```{r}

# Model 4: Retired ~ Age + Income + Gender + Education
model4 <- glm(Retired ~ Age + Income + Gender + Education, data = data, family = binomial)

# Summary of the model
summary(model4)

```