### 2024-08-02 presentation-exercise/presentation-exercise.qmd

---
title: "Presentation Exercise"
---

Placeholder file for the future data/results presentation exercise.

# load libraries

```{r}

library(ggplot2)
library(tidyverse)
library(here)
library(plotly)
library(flextable)

```

# Import data

```{r}

data <- read.csv("piedata.csv", header = TRUE, sep = ",")

data$Category <- as.factor(data$Category)
data$Value <- as.numeric(data$Value)

str(data)

```

# Create the Graph

```{r}

blue_palette <- scales::brewer_pal(palette = "Blues")(4)

ggplot(data, aes(x = "", y = Value, fill = Category)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  scale_fill_manual(values = blue_palette) +  # Use the blue color palette
  labs(title = "RESIDENCES ARE THE LEADING LOCATION OF GUN THEFTS FROM CARS") +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5, margin = margin(b = 20)))  # Center title and adjust bottom margin

```

```{r}
set_flextable_defaults(
  font.color = "black",
  border.color = "lightblue",
  theme_fun = "theme_box")

dat <- data.frame(
  Category = c("Other", "Parking Lot", "Residence", "Road"),
  Value = c(12, 32, 36, 20 )

)
flextable(dat)

```

```{r}

init_flextable_defaults()
flextable(dat)
```