### 2024-08-02 tidytuesday-exercise/tidytuesday-exercise(draft).qmd

---
title: "Tidy Tuesday Exercise"
---

Placeholder file for the future Tidy Tuesday exercise.

```{r}
library(tidyverse)
library(ggplot2)
library(knitr)
library(dplyr)
library(here)
```

```{r}
#correctly removed first row and replaced with correct labels or variable names
Data <- read.csv("C:/Users/Leonel/Desktop/MSDA/MS Data Analytics/Current Class/DA 6833/Practicum 2 Github/leonelsalazar-P2-portfolio/tidytuesday-exercise/finalists.csv", header = T, na.strings = "?",
stringsAsFactors = T)
```

```{r}
str(Data)
```

```{r}
view(Data)
```

```{r}
# Select all columns except 4
contestants_data <- dplyr::select(Data, -c(4,5))
```

```{r}
contestants_data <- contestants_data %>%
   mutate(Birthday = as.Date(Birthday, format = "%d-%b-%y")) %>%
   mutate(Birthday = format(Birthday, "%d-%m-%y"))
```

```{r}
view(contestants_data)
```

```{r}
str(contestants_data)
```

```{r}
# Convert Contestant to character to clean it
contestants_data$Contestant <- as.character(contestants_data$Contestant)

# Convert Contestant column to UTF-8 encoding to handle encoding issues
contestants_data$Contestant <- iconv(contestants_data$Contestant, to = "UTF-8")

# Remove non-printable characters and unwanted symbols
contestants_data$Contestant <- gsub("[^[:print:]]", "", contestants_data$Contestant)

# Remove specific unwanted characters (e.g., quotes and slashes) if needed
contestants_data$Contestant <- gsub("[\"/]", "", contestants_data$Contestant)

# Convert Contestant back to factor
contestants_data$Contestant <- as.factor(contestants_data$Contestant)

# Ensure Birthday is treated as a factor (if needed)
contestants_data$Birthday <- as.factor(contestants_data$Birthday)

# Remove rows with NA values in Contestant, Birthday, and Season columns (if needed)
contestants_data_clean <- contestants_data %>%
  drop_na(Contestant, Birthday, Season)
```

```{r}
str(contestants_data_clean)
```
```{r}
#contestants_data_clean$Contestant <- as.numeric(contestants_data_clean$Contestant)
```

```{r}
view(contestants_data_clean)
```

```{r}
contestants_data_clean <- contestants_data_clean %>%
  drop_na(Birthplace)

```

```{r}
View(contestants_data_clean)
```

```{r}
write.csv(contestants_data_clean, "C:/Users/Leonel/Desktop/MSDA/MS Data Analytics/Current Class/DA 6833/Practicum 2 Github/leonelsalazar-P2-portfolio/tidytuesday-exercise/contestants_data_clean.csv", row.names = FALSE)
```

```{r}
library(ggplot2)

# Convert Contestant to numeric (just for plotting)
contestants_data_clean$Contestant <- as.numeric(contestants_data_clean$Contestant)

# Create histogram of Contestant numbers
ggplot(contestants_data_clean, aes(x = Contestant)) +
  geom_histogram(bins = 20, fill = "steelblue", color = "black") +
  labs(title = "Distribution of Contestants",
       x = "Contestant (Numeric Representation)",
       y = "Count") +
  theme_minimal()
```

```{r}
contestants_data_clean$Contestant_Num <- as.numeric(contestants_data_clean$Contestant)

ggplot(contestants_data_clean, aes(x = factor(Season), y = Contestant_Num)) +
  geom_boxplot(fill = "lightblue", color = "black") +
  labs(title = "Box Plot of Contestants by Season",
       x = "Season",
       y = "Contestant (Numeric Representation)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```