### 2024-12-03 R Projects/Bank_Marketing_Case_Study.qmd
updated assignments to github

---
title: "Bank Marketing Case Study"
author: "Collin Real, Leonel Salazar, Seth Harris, Joaquin Ramirez"
format: docx
---

---
title: "Bank Marketing Case Study"

format: docx
---

```{r}
library(tidyverse)
library(ggplot2)
library(here)
library(caret)
library(dplyr)
library(stats)
library(car)
library(pscl)
library(flextable)
library(corrplot)
library(pROC)
library(randomForest)

```

```{r}

# Use the here function to construct the file path and import the dataset
data <- read.csv(here("bank-additional_clean.csv"), header = TRUE, sep = ",")

# View the first few rows of the dataset
head(data)

str(data)

```

```{r}

# Function to check for missing (NA) values and "unknown" entries in all columns
check_missing_unknown <- function(data) {
  result <- data.frame(
    Variable = colnames(data),
    Missing_Count = sapply(data, function(x) sum(is.na(x))),  # Count of NA values
    Missing_Percentage = sapply(data, function(x) mean(is.na(x)) * 100),  # Percentage of NA values
    Unknown_Count = sapply(data, function(x) sum(tolower(as.character(x)) == "unknown", na.rm = TRUE)),  # Count of "unknown"
    Unknown_Percentage = sapply(data, function(x) mean(tolower(as.character(x)) == "unknown", na.rm = TRUE) * 100)  # Percentage of "unknown"
  )
  return(result)
}

missing_unknown_summary <- check_missing_unknown(data)

# Display the summary table
print(missing_unknown_summary)

```

```{r}

# Convert all integer and numeric variables to numeric type
data[] <- lapply(data, function(x) {
  if (is.integer(x) || is.numeric(x)) {
    return(as.numeric(x))  # Convert to numeric
  } else if (is.character(x)) {
    return(factor(x))  # Convert character to factor
  } else {
    return(x)  # Leave other types unchanged
  }
})

# Verify the changes
str(data)

```

```{r}

# Transform the 'pdays' column
data_clean <- data %>%
  mutate(pdays = ifelse(pdays == 999, "not previously contacted", "previously contacted"))

# Convert 'pdays' to a factor
data_clean$pdays <- as.factor(data_clean$pdays)

# Display the first few rows to verify the change
head(data_clean)

```

```{r}

# Remove the 'duration' variable
data_clean <- select(data_clean, -duration)

# Save the cleaned dataset as a new CSV file
write.csv(data_clean, "data_clean.csv", row.names = FALSE)

# Verify the dataset has been saved
file.exists("data_clean.csv")

```

```{r}

# Export the dataset to a CSV file in the same location
write_csv(data, here("data_new.csv"))

# Confirm the file is saved by showing the file path
cat("Data has been exported to:", here("data_new.csv"))

```

```{r}

# Check the distribution of the target variable 'y'
table(data$y)

# Calculate the percentage distribution of 'y'
prop.table(table(data$y)) * 100

```

```{r}

# Function to count "unknown" values in each column
count_unknowns <- function(df) {
  sapply(df, function(x) sum(x == "unknown", na.rm = TRUE))
}

# Apply the function to the dataset
unknown_counts <- count_unknowns(data)

# Display the counts
print(unknown_counts)

# Calculate the total number of "unknown" values across all variables
total_unknowns <- sum(unknown_counts)
print(paste("Total number of 'unknown' values across all variables:", total_unknowns))

```

```{r}

# Check for NA values in each column
na_counts <- sapply(data, function(x) sum(is.na(x)))

# Display the counts of NA values for each column
print(na_counts)

# Calculate the total number of NA values across all columns
total_na <- sum(na_counts)
print(paste("Total number of NA values across all columns:", total_na))

```

```{r}

# Up-sample the minority class
set.seed(123)  # For reproducibility
data_clean_upsampled <- upSample(x = data_clean %>% select(-y), y = data_clean$y)  # 'y' is the target variable

# Check the new class distribution
table(data_clean_upsampled$Class)

```
```{r}

# Down-sample the majority class
set.seed(123)  # For reproducibility
data_clean_downsampled <- downSample(x = data_clean %>% select(-y), y = data_clean$y)  # 'y' is the target variable

# Check the new class distribution
table(data_clean_downsampled$Class)

```

```{r}

# Load necessary libraries
library(dplyr)

# Calculate the count of "unknown" values for each column
unknown_counts <- sapply(data_clean_downsampled, function(x) sum(x == "unknown", na.rm = TRUE))

# Convert the counts to a data frame for better visualization
unknown_counts_df <- data.frame(
  Variable = names(unknown_counts),
  Unknown_Count = unknown_counts
)

# Calculate the proportion of "unknown" values for each column
unknown_counts_df <- unknown_counts_df %>%
  mutate(Total_Count = nrow(data_clean_downsampled),  # Total rows in the dataset
         Unknown_Proportion = Unknown_Count / Total_Count * 100)  # Proportion in percentage

# Display the data frame with counts and proportions
print(unknown_counts_df)

# Calculate the overall percentage of "unknown" values across all variables
total_unknowns <- sum(unknown_counts)
total_values <- nrow(data_clean_downsampled) * ncol(data_clean_downsampled)  # Total number of data points
overall_unknown_percentage <- (total_unknowns / total_values) * 100

print(paste("Overall percentage of 'unknown' values in the dataset:", round(overall_unknown_percentage, 2), "%"))

```

```{r}

# Remove rows with 'unknown' values
data_clean_downsampled_no_unknown <- data_clean_downsampled %>%
  filter_all(~ . != "unknown")

# Check the new size of the dataset
print(dim(data_clean_downsampled_no_unknown))

```

```{r}

# Load necessary libraries
library(dplyr)

# Remove rows with 'unknown' values from the downsampled data
data_clean_downsampled_no_unknown <- data_clean_downsampled %>%
  filter_all(~ . != "unknown")

# Check the distribution of the target variable 'y' after removing unknowns
balance_after_cleaning <- table(data_clean_downsampled_no_unknown$Class)  # Assuming 'Class' is the name of the target variable column

# Print the class balance
print(balance_after_cleaning)

# Calculate and display the proportion of each class
balance_proportion <- prop.table(balance_after_cleaning) * 100
print(balance_proportion)

```

```{r}

# Export the dataset to a CSV file in the same location
write_csv(data, here("data_clean_downsampled_no_unknown .csv"))

# Confirm the file is saved by showing the file path
cat("Data has been exported to:", here("data_clean_downsampled_no_unknown .csv"))
```

```{r}

df <- data_clean_downsampled_no_unknown

# Rename the column 'Class' to 'y'
df <- df %>% rename(y = Class)

# Identify categorical and numeric variables
variables <- names(df)
var_types <- ifelse(sapply(df, is.numeric), "Numeric", "Categorical")

# Create a data frame to store variable names and their types
var_table <- data.frame(Variable = variables, Type = var_types)

# Create a flextable
flex_table <- flextable(var_table)

# Apply custom formatting:
# Highlight "Categorical" types with light blue, and "Numeric" types with light pink
flex_table <- flextable::bg(flex_table, j = "Type", i = ~ Type == "Categorical", bg = "lightblue")  # Highlight categorical
flex_table <- flextable::bg(flex_table, j = "Type", i = ~ Type == "Numeric", bg = "lightpink")      # Highlight numeric

# Adjust column widths for better readability
flex_table <- autofit(flex_table)

# Display the flextable
flex_table

```

```{r}

# Set up a 3x3 plotting area
par(mfrow=c(3,3))

# Identify categorical variables (factor or character type)
categorical_vars <- sapply(df, function(x) is.factor(x) | is.character(x))  
categorical_data <- df[, categorical_vars]  # Subset the dataframe for categorical variables

# Loop through all categorical variables and plot bar plots
for (var in names(categorical_data)) {
  barplot(table(categorical_data[[var]]), main=paste("Bar Plot of", var), xlab=var, col="lightblue")
}

# Reset the plotting layout to 1x1 after plotting
par(mfrow=c(1,1))

```

```{r}

# Set up a 3x3 plotting area
par(mfrow=c(3,3))

# Loop through all the columns and plot histograms for continuous (numeric) variables
numeric_vars <- sapply(df, is.numeric)  # Identify numeric variables
continuous_vars <- df[, numeric_vars]   # Subset the dataframe for numeric variables

# Plot histograms for each numeric variable
for (var in names(continuous_vars)) {
  hist(continuous_vars[[var]], main=paste("Histogram of", var), xlab=var, col="lightblue", breaks=20)
}

# Reset the plotting layout to 1x1 after plotting
par(mfrow=c(1,1))

```

```{r}

# Set up a 3x3 plotting area
par(mfrow=c(1,2))

# Identify numeric variables in the dataset
numeric_vars <- sapply(df, is.numeric)

# Subset the dataset to include only the numeric variables
numeric_data <- df[, numeric_vars]

# Loop through the numeric variables and create boxplots
for (var in names(numeric_data)) {
  boxplot(numeric_data[[var]], main=paste("Boxplot of", var), xlab=var, col="lightblue")
}

# Reset the plotting layout to 1x1 after plotting

```

```{r}

library(corrplot)

# Select only the numeric variables from the dataset
numeric_vars <- df[, sapply(df, is.numeric)]

# Create the correlation matrix
cor_matrix <- cor(numeric_vars, use="complete.obs")

# Visualize the correlation matrix
corrplot(cor_matrix, method="circle", type="lower", 
         tl.col="black", tl.cex=0.8, title="Correlation Matrix of Numeric Variables",
         mar=c(0,0,1,0))

```
### Key Assumptions of Logistic Regression:

1. **Binary Outcome Variable**: The dependent variable should be binary.
2. **Independence of Observations**: Observations should be independent of each other.
3. **No Multicollinearity**: Predictor variables should not be highly correlated with each other.
4. **Linearity of Independent Variables and Log-Odds**: There should be a linear relationship between continuous predictors and the log-odds of the outcome.
5. **Sufficient Sample Size**: Logistic regression requires a large sample size to provide reliable results.

### Steps to Test the Assumptions

### 1. **Check for Binary Outcome Variable**
Ensure the dependent variable (`y`) is binary.

```{r}
# Check the levels of the outcome variable
table(df$y)
```

```{r}

# Check for variables with only one level
lapply(df, function(x) length(unique(x)))

```

```{r}

# Set a seed for reproducibility
set.seed(123)

# Split the data into training (70%) and testing (30%) sets
trainIndex <- createDataPartition(df$y, p = 0.7, list = FALSE)
train_data <- df[trainIndex, ]  # 70% training data
test_data  <- df[-trainIndex, ] # 30% test data

```

```{r}

# Fit the logistic regression model on the training set
logit_model_train <- glm(y ~ age + job + marital + education + housing + loan + contact + month + 
                         day_of_week + campaign + pdays + previous + poutcome + emp.var.rate + 
                         cons.price.idx + cons.conf.idx, 
                         family = binomial, data = train_data)

# Display the summary of the model
summary(logit_model_train)

```

- **Interpretation**: Variables with VIF > 5 may have multicollinearity issues.

```{r}

# Calculate VIF values for the logistic regression model
vif_values_train <- vif(logit_model_train)

# Print the VIF values
print(vif_values_train)

```

```{r}

# Example: Remove a variable with high VIF and refit the model
logit_model_train_refit <- glm(y ~ age + job + marital + education + housing + loan + contact + month + 
                               day_of_week + campaign + pdays + previous + poutcome + emp.var.rate + 
                               cons.price.idx, 
                               family = binomial, data = train_data)

# Recheck VIF values for the refit model
vif_values_refit <- vif(logit_model_train_refit)
print(vif_values_refit)

```

```{r}
# Check the number of events (e.g., 1's and 0's in the outcome variable)
table(train_data$y)

# Ensure the number of events is at least 10 times the number of predictors
```

```{r}

# Predict probabilities for the training data using the logistic regression model
predicted_probabilities_train <- predict(logit_model_train, newdata = train_data, type = "response")

# Convert probabilities to binary outcome (using 0.5 as the cutoff)
predicted_classes_train <- ifelse(predicted_probabilities_train > 0.5, 1, 0)

# Create confusion matrix for training data
confusion_matrix_train <- table(predicted_classes_train, train_data$y)
print(confusion_matrix_train)

# Extract the values from the confusion matrix
TN <- confusion_matrix_train[1,1]  # True Negatives
FP <- confusion_matrix_train[1,2]  # False Positives
FN <- confusion_matrix_train[2,1]  # False Negatives
TP <- confusion_matrix_train[2,2]  # True Positives

# Calculate accuracy
accuracy_train <- (TP + TN) / sum(confusion_matrix_train)

# Calculate precision, recall, and F1 score
precision_train <- ifelse((TP + FP) > 0, TP / (TP + FP), 0)  # TP / (TP + FP)
recall_train <- ifelse((TP + FN) > 0, TP / (TP + FN), 0)     # TP / (TP + FN)
f1_score_train <- ifelse((precision_train + recall_train) > 0, 
                         2 * ((precision_train * recall_train) / (precision_train + recall_train)), 
                         0)

# Print the performance metrics for the training data
print(paste("Accuracy:", accuracy_train))
print(paste("Precision:", precision_train))
print(paste("Recall:", recall_train))
print(paste("F1 Score:", f1_score_train))

```

```{r}

accuracy <- 0.7722   
precision <- 0.7259  
recall <- 0.8174     
f1_score <- 0.7689

# Create a dataframe with the performance metrics
metrics_data <- data.frame(
  Metric = c("Accuracy", "Precision", "Recall", "F1 Score"),
  Value = c(accuracy, precision, recall, f1_score)
)

# Create the flextable
performance_table <- flextable(metrics_data)

# Apply consistent styling to the flextable with alternating colors
performance_table <- performance_table %>%
  color(j = 1, color = "black") %>%                        # Text color for Metric column
  color(j = 2, color = "darkblue") %>%                     # Text color for Value column
  bg(part = "body", bg = "lightgray", i = ~ Metric == "Accuracy") %>% # Background color for Accuracy
  bg(part = "body", bg = "lightblue", i = ~ Metric == "Precision") %>% # Background color for Precision
  bg(part = "body", bg = "lightgray", i = ~ Metric == "Recall") %>%   # Background color for Recall
  bg(part = "body", bg = "lightblue", i = ~ Metric == "F1 Score") %>% # Background color for F1 Score
  align(j = 2, align = "center", part = "body") %>%        # Center-align the values
  autofit()                                                # Adjust column widths

# Print the flextable
performance_table

```

```{r}

# Predict probabilities for the training set using the logistic regression model
predicted_probabilities_train <- predict(logit_model_train, newdata = train_data, type = "response")

# Create the ROC curve using the training data
roc_curve_train <- roc(train_data$y, predicted_probabilities_train)

# Plot the ROC curve for the training data
plot(roc_curve_train, main = "ROC Curve for Logistic Regression Model (Training Data)", col = "blue", lwd = 2)

# Calculate the AUC for the training data
auc_train <- auc(roc_curve_train)

# Print the AUC value for the training data
print(paste("AUC for Training Data:", auc_train))

```

```{r}

# Predict probabilities for the test data using the logistic regression model
predicted_probabilities_test <- predict(logit_model_train, newdata = test_data, type = "response")

# Convert probabilities to binary outcomes (using 0.5 as the cutoff)
predicted_classes_test <- ifelse(predicted_probabilities_test > 0.5, 1, 0)

# Create confusion matrix for the test data
confusion_matrix_test <- table(predicted_classes_test, test_data$y)
print(confusion_matrix_test)

# Extract values from the confusion matrix for test data
TN_test <- confusion_matrix_test[1,1]  # True Negatives
FP_test <- confusion_matrix_test[1,2]  # False Positives
FN_test <- confusion_matrix_test[2,1]  # False Negatives
TP_test <- confusion_matrix_test[2,2]  # True Positives

# Calculate accuracy for the test data
accuracy_test <- (TP_test + TN_test) / sum(confusion_matrix_test)

# Calculate precision, recall, and F1 score for the test data
precision_test <- ifelse((TP_test + FP_test) > 0, TP_test / (TP_test + FP_test), 0)
recall_test <- ifelse((TP_test + FN_test) > 0, TP_test / (TP_test + FN_test), 0)
f1_score_test <- ifelse((precision_test + recall_test) > 0, 
                        2 * ((precision_test * recall_test) / (precision_test + recall_test)), 
                        0)

# Print the performance metrics for the test data
print(paste("Accuracy (Test):", accuracy_test))
print(paste("Precision (Test):", precision_test))
print(paste("Recall (Test):", recall_test))
print(paste("F1 Score (Test):", f1_score_test))

```

```{r}

# Checking for multicollinearity (VIF values) - same as training data, but for the model
vif_values_test <- vif(logit_model_train)  # VIF is the same as in training
print(vif_values_test)

```

```{r}

# Create the ROC curve for the test data
roc_curve_test <- roc(test_data$y, predicted_probabilities_test)

# Plot the ROC curve for the test data
plot(roc_curve_test, main = "ROC Curve for Logistic Regression Model (Test Data)", col = "blue", lwd = 2)

# Calculate the AUC for the test data
auc_test <- auc(roc_curve_test)

# Print the AUC value for the test data
print(paste("AUC for Test Data:", auc_test))

```
```{r}

# Create a dataframe for comparison
metrics_comparison <- data.frame(
  Metric = c("Accuracy", "Precision", "Recall", "F1 Score", "AUC"),
  Training = c(0.7721, 0.7259, 0.8174, 0.7689, 0.8531),
  Test = c(0.7406, 0.6846, 0.7917, 0.7343, 0.8183)
)

# Create the flextable
comparison_table <- flextable(metrics_comparison)

# Format the flextable with some custom styles
comparison_table <- comparison_table %>%
  color(j = 1, color = "black") %>%                        # Text color for Metric column
  color(j = 2:3, color = "darkblue") %>%                   # Text color for values in Training and Test columns
  bg(part = "body", bg = "lightgray", i = ~ Metric == "Accuracy") %>%
  bg(part = "body", bg = "lightblue", i = ~ Metric == "Precision") %>%
  bg(part = "body", bg = "lightgray", i = ~ Metric == "Recall") %>%
  bg(part = "body", bg = "lightblue", i = ~ Metric == "F1 Score") %>%
  bg(part = "body", bg = "lightgray", i = ~ Metric == "AUC") %>%
  align(j = 2:3, align = "center", part = "body") %>%      # Center-align the values
  autofit()                                                # Adjust column widths

# Print the flextable
comparison_table

```

```{r}

# Fit the Random Forest model
set.seed(123)  # Set a seed for reproducibility

rf_model <- randomForest(y ~ age + job + marital + education + housing + loan + contact + month + 
                         day_of_week + campaign + pdays + previous + poutcome + emp.var.rate + 
                         cons.price.idx + cons.conf.idx, 
                         data = df, ntree = 500, mtry = 3, importance = TRUE)

# Print the model summary
print(rf_model)

```

```{r}

# Predict the class (0/1) for the test data
predicted_rf <- predict(rf_model, df)

# Create confusion matrix
conf_matrix_rf <- table(predicted_rf, df$y)

# Print confusion matrix
print(conf_matrix_rf)

# Calculate accuracy
accuracy_rf <- mean(predicted_rf == df$y)
print(paste("Random Forest Accuracy:", accuracy_rf))

```

```{r}

# Get predicted probabilities from the Random Forest model
predicted_probabilities_rf <- predict(rf_model, df, type = "prob")[,2]

# Plot the ROC curve
roc_curve_rf <- roc(df$y, predicted_probabilities_rf)
plot(roc_curve_rf)

# Calculate the AUC
auc_rf <- auc(roc_curve_rf)
print(paste("AUC for Random Forest:", auc_rf))

```
```{r}

# Get variable importance
importance_rf <- importance(rf_model)

# Plot variable importance
varImpPlot(rf_model)

```

```{r}

# Set a seed for reproducibility
set.seed(123)

# Train the Random Forest model on the training data
rf_model_train <- randomForest(y ~ age + job + marital + education + housing + loan + contact + month +
                               day_of_week + campaign + pdays + previous + poutcome + emp.var.rate +
                               cons.price.idx + cons.conf.idx,
                               data = train_data, ntree = 500, mtry = 3, importance = TRUE)

# Print the summary of the Random Forest model (training data)
print(rf_model_train)

```

```{r}

# Manually input the confusion matrix values
TN_rf_train <- 177  # True Negatives
FP_rf_train <- 60   # False Positives
FN_rf_train <- 85   # False Negatives
TP_rf_train <- 174  # True Positives

# Calculate accuracy
accuracy_rf_train <- (TP_rf_train + TN_rf_train) / (TP_rf_train + TN_rf_train + FP_rf_train + FN_rf_train)
print(paste("Accuracy (Training):", accuracy_rf_train))

# Calculate precision, recall, and F1 score, handling division by zero
precision_rf_train <- ifelse((TP_rf_train + FP_rf_train) > 0, TP_rf_train / (TP_rf_train + FP_rf_train), 0)
recall_rf_train <- ifelse((TP_rf_train + FN_rf_train) > 0, TP_rf_train / (TP_rf_train + FN_rf_train), 0)
f1_score_rf_train <- ifelse((precision_rf_train + recall_rf_train) > 0,
                            2 * ((precision_rf_train * recall_rf_train) / (precision_rf_train + recall_rf_train)),
                            0)

# Print the metrics for the training data
print(paste("Precision (Training):", precision_rf_train))
print(paste("Recall (Training):", recall_rf_train))
print(paste("F1 Score (Training):", f1_score_rf_train))

```
```{r}

# Manually input the confusion matrix values for test data
TN_rf_test <- 82  # True Negatives
FP_rf_test <- 40  # False Positives
FN_rf_test <- 19  # False Negatives
TP_rf_test <- 71  # True Positives

# Calculate accuracy for test data
accuracy_rf_test <- (TP_rf_test + TN_rf_test) / (TP_rf_test + TN_rf_test + FP_rf_test + FN_rf_test)
print(paste("Accuracy (Test):", accuracy_rf_test))

# Calculate precision, recall, and F1 score, handling division by zero
precision_rf_test <- ifelse((TP_rf_test + FP_rf_test) > 0, TP_rf_test / (TP_rf_test + FP_rf_test), 0)
recall_rf_test <- ifelse((TP_rf_test + FN_rf_test) > 0, TP_rf_test / (TP_rf_test + FN_rf_test), 0)
f1_score_rf_test <- ifelse((precision_rf_test + recall_rf_test) > 0,
                           2 * ((precision_rf_test * recall_rf_test) / (precision_rf_test + recall_rf_test)),
                           0)

# Print the metrics for the test data
print(paste("Precision (Test):", precision_rf_test))
print(paste("Recall (Test):", recall_rf_test))
print(paste("F1 Score (Test):", f1_score_rf_test))

```

```{r}

# Predict probabilities for the ROC and AUC on the training data
predicted_probabilities_rf_train <- predict(rf_model_train, newdata = train_data, type = "prob")[, 2]
roc_curve_rf_train <- roc(train_data$y, predicted_probabilities_rf_train)
auc_rf_train <- auc(roc_curve_rf_train)

# Predict probabilities for the ROC and AUC on the test data
predicted_probabilities_rf_test <- predict(rf_model_train, newdata = test_data, type = "prob")[, 2]
roc_curve_rf_test <- roc(test_data$y, predicted_probabilities_rf_test)
auc_rf_test <- auc(roc_curve_rf_test)

# Print AUC for both datasets
print(paste("AUC (Training):", auc_rf_train))
print(paste("AUC (Test):", auc_rf_test))

```

```{r}

# Create a dataframe for Random Forest comparison
rf_metrics_comparison <- data.frame(
  Metric = c("Accuracy", "Precision", "Recall", "F1 Score", "AUC"),
  Training = c(accuracy_rf_train, precision_rf_train, recall_rf_train, f1_score_rf_train, auc_rf_train),
  Test = c(accuracy_rf_test, precision_rf_test, recall_rf_test, f1_score_rf_test, auc_rf_test)
)

# Create the flextable for Random Forest model comparison
rf_comparison_table <- flextable(rf_metrics_comparison)

# Format the flextable with some custom styles
rf_comparison_table <- rf_comparison_table %>%
  color(j = 1, color = "black") %>%
  color(j = 2:3, color = "darkblue") %>%
  bg(part = "body", bg = "lightgray", i = ~ Metric == "Accuracy") %>%
  bg(part = "body", bg = "lightblue", i = ~ Metric == "Precision") %>%
  bg(part = "body", bg = "lightgray", i = ~ Metric == "Recall") %>%
  bg(part = "body", bg = "lightblue", i = ~ Metric == "F1 Score") %>%
  bg(part = "body", bg = "lightgray", i = ~ Metric == "AUC") %>%
  align(j = 2:3, align = "center", part = "body") %>%
  autofit()

# Print the flextable
rf_comparison_table

```

```{r}

# Set up 10-fold cross-validation
train_control <- trainControl(method = "cv", number = 10)

# Train the logistic regression model using 10-fold cross-validation
logit_model_cv <- train(y ~ age + job + marital + education + housing + loan + contact + month + 
                        day_of_week + campaign + pdays + previous + poutcome + emp.var.rate + 
                        cons.price.idx + cons.conf.idx,
                        data = df, method = "glm", family = binomial,
                        trControl = train_control)

# Print the results of the cross-validation
print(logit_model_cv)

```

```{r}

# Set up 10-fold cross-validation
train_control_rf <- trainControl(method = "cv", number = 10)

# Train the Random Forest model using 10-fold cross-validation on the training data
rf_model_cv <- train(y ~ age + job + marital + education + housing + loan + contact + month + 
                     day_of_week + campaign + pdays + previous + poutcome + emp.var.rate + 
                     cons.price.idx + cons.conf.idx,
                     data = train_data, method = "rf", 
                     ntree = 500,  # Set the number of trees
                     trControl = train_control_rf, importance = TRUE)

# Print the results of the cross-validation
print(rf_model_cv)

```
```{r}

# Set up 10-fold cross-validation for the test data
train_control_test <- trainControl(method = "cv", number = 10)

# Perform logistic regression with 10-fold cross-validation on the test data
logit_model_cv_test <- train(y ~ age + job + marital + education + housing + loan + contact + month + 
                             day_of_week + campaign + pdays + previous + poutcome + emp.var.rate + 
                             cons.price.idx + cons.conf.idx,
                             data = test_data, method = "glm", family = binomial,
                             trControl = train_control_test)

# Print the cross-validation results for the test data
print(logit_model_cv_test)

# Predict the probabilities on the test data
predicted_probabilities_test_cv <- predict(logit_model_cv_test, newdata = test_data, type = "prob")[, 2]

# Convert probabilities to binary outcome (using 0.5 as the cutoff)
predicted_classes_test_cv <- ifelse(predicted_probabilities_test_cv > 0.5, 1, 0)

# Create a confusion matrix for the test data
confusion_matrix_test_cv <- table(predicted_classes_test_cv, test_data$y)
print(confusion_matrix_test_cv)

# Manually input the confusion matrix values
TN_test_cv <- 82  # True Negatives
FP_test_cv <- 24  # False Positives
FN_test_cv <- 19  # False Negatives
TP_test_cv <- 87  # True Positives

# Calculate accuracy for the test data
accuracy_test_cv <- (TP_test_cv + TN_test_cv) / (TP_test_cv + TN_test_cv + FP_test_cv + FN_test_cv)
print(paste("Accuracy (Test with CV):", accuracy_test_cv))

# Calculate precision, recall, and F1 score, handling division by zero
precision_test_cv <- ifelse((TP_test_cv + FP_test_cv) > 0, TP_test_cv / (TP_test_cv + FP_test_cv), 0)
recall_test_cv <- ifelse((TP_test_cv + FN_test_cv) > 0, TP_test_cv / (TP_test_cv + FN_test_cv), 0)
f1_score_test_cv <- ifelse((precision_test_cv + recall_test_cv) > 0,
                           2 * ((precision_test_cv * recall_test_cv) / (precision_test_cv + recall_test_cv)),
                           0)

# Print the metrics for the test data
print(paste("Precision (Test with CV):", precision_test_cv))
print(paste("Recall (Test with CV):", recall_test_cv))
print(paste("F1 Score (Test with CV):", f1_score_test_cv))

```

```{r}

# Set up 10-fold cross-validation for the test data
train_control_rf_test <- trainControl(method = "cv", number = 10)

# Perform Random Forest with 10-fold cross-validation on the test data
rf_model_cv_test <- train(y ~ age + job + marital + education + housing + loan + contact + month + 
                          day_of_week + campaign + pdays + previous + poutcome + emp.var.rate + 
                          cons.price.idx + cons.conf.idx,
                          data = test_data, method = "rf", 
                          ntree = 500,  # Number of trees in the Random Forest
                          trControl = train_control_rf_test)

# Print the cross-validation results for the test data
print(rf_model_cv_test)

# Predict the classes on the test data using the cross-validated Random Forest model
predicted_rf_test_cv <- predict(rf_model_cv_test, newdata = test_data)

# Create a confusion matrix for the test data
confusion_matrix_rf_test_cv <- table(predicted_rf_test_cv, test_data$y)
print(confusion_matrix_rf_test_cv)

# Manually input the confusion matrix values
TN_rf_test_cv <- 91  # True Negatives
FP_rf_test_cv <- 26  # False Positives
FN_rf_test_cv <- 10  # False Negatives
TP_rf_test_cv <- 85  # True Positives

# Calculate accuracy for the test data
accuracy_rf_test_cv <- (TP_rf_test_cv + TN_rf_test_cv) / (TP_rf_test_cv + TN_rf_test_cv + FP_rf_test_cv + FN_rf_test_cv)
print(paste("Accuracy (Test with CV - Random Forest):", accuracy_rf_test_cv))

# Calculate precision, recall, and F1 score, handling division by zero
precision_rf_test_cv <- ifelse((TP_rf_test_cv + FP_rf_test_cv) > 0, TP_rf_test_cv / (TP_rf_test_cv + FP_rf_test_cv), 0)
recall_rf_test_cv <- ifelse((TP_rf_test_cv + FN_rf_test_cv) > 0, TP_rf_test_cv / (TP_rf_test_cv + FN_rf_test_cv), 0)
f1_score_rf_test_cv <- ifelse((precision_rf_test_cv + recall_rf_test_cv) > 0,
                              2 * ((precision_rf_test_cv * recall_rf_test_cv) / (precision_rf_test_cv + recall_rf_test_cv)),
                              0)

# Print the metrics for the test data
print(paste("Precision (Test with CV - Random Forest):", precision_rf_test_cv))
print(paste("Recall (Test with CV - Random Forest):", recall_rf_test_cv))
print(paste("F1 Score (Test with CV - Random Forest):", f1_score_rf_test_cv))

```
```{r}

# ---------------------- Logistic Regression with CV ----------------------

# Set up 10-fold cross-validation for the logistic regression on the test data
train_control_logit_test <- trainControl(method = "cv", number = 10)

# Perform logistic regression with 10-fold cross-validation on the test data
logit_model_cv_test <- train(y ~ age + job + marital + education + housing + loan + contact + month + 
                             day_of_week + campaign + pdays + previous + poutcome + emp.var.rate + 
                             cons.price.idx + cons.conf.idx,
                             data = test_data, method = "glm", family = binomial,
                             trControl = train_control_logit_test)

# Predict probabilities for ROC curve on the test data
predicted_probabilities_logit_cv_test <- predict(logit_model_cv_test, newdata = test_data, type = "prob")[, 2]

# Calculate the ROC curve and AUC for logistic regression on the test data
roc_logit_test <- roc(test_data$y, predicted_probabilities_logit_cv_test)
auc_logit_test <- auc(roc_logit_test)

# Print the AUC for logistic regression on the test data
print(paste("AUC (Logistic Regression with CV - Test Data):", auc_logit_test))

# Plot the ROC curve for logistic regression on the test data
plot(roc_logit_test, main = "ROC Curve - Logistic Regression (Test Data)", col = "blue")

# ---------------------- Random Forest with CV ----------------------

# Set up 10-fold cross-validation for the random forest on the test data
train_control_rf_test <- trainControl(method = "cv", number = 10)

# Perform Random Forest with 10-fold cross-validation on the test data
rf_model_cv_test <- train(y ~ age + job + marital + education + housing + loan + contact + month + 
                          day_of_week + campaign + pdays + previous + poutcome + emp.var.rate + 
                          cons.price.idx + cons.conf.idx,
                          data = test_data, method = "rf", 
                          ntree = 500,  # Number of trees in the Random Forest
                          trControl = train_control_rf_test)

# Predict probabilities for ROC curve on the test data
predicted_probabilities_rf_cv_test <- predict(rf_model_cv_test, newdata = test_data, type = "prob")[, 2]

# Calculate the ROC curve and AUC for Random Forest on the test data
roc_rf_test <- roc(test_data$y, predicted_probabilities_rf_cv_test)
auc_rf_test <- auc(roc_rf_test)

# Print the AUC for Random Forest on the test data
print(paste("AUC (Random Forest with CV - Test Data):", auc_rf_test))

# Plot the ROC curve for Random Forest on the test data
plot(roc_rf_test, main = "ROC Curve - Random Forest (Test Data)", col = "red")

```
```{r}

# Logistic Regression Cross-Validation Results (updated from the images)
accuracy_logit <- 0.7972  # Accuracy
precision_logit <- 0.7384  # Precision
recall_logit <- 0.8208  # Recall
f1_score_logit <- 0.8018  # F1 Score
auc_logit_test <- 0.8907  # AUC for logistic regression

# Random Forest Cross-Validation Results (updated from the images)
accuracy_rf <- 0.8302  # Accuracy
precision_rf <- 0.7658  # Precision
recall_rf <- 0.8947  # Recall
f1_score_rf <- 0.8254  # F1 Score
auc_rf_test <- 0.9395  # AUC for random forest

# Create a dataframe for comparison
metrics_comparison <- data.frame(
  Metric = c("Accuracy", "Precision", "Recall", "F1 Score", "AUC"),
  Logistic_Regression = c(accuracy_logit, precision_logit, recall_logit, f1_score_logit, auc_logit_test),
  Random_Forest = c(accuracy_rf, precision_rf, recall_rf, f1_score_rf, auc_rf_test)
)

# Create the flextable for model comparison
comparison_table <- flextable(metrics_comparison)

# Format the flextable with custom styles
comparison_table <- comparison_table %>%
  bg(part = "body", bg = "lightgray", i = ~ Metric == "Accuracy") %>%
  bg(part = "body", bg = "lightblue", i = ~ Metric == "Precision") %>%
  bg(part = "body", bg = "lightgray", i = ~ Metric == "Recall") %>%
  bg(part = "body", bg = "lightblue", i = ~ Metric == "F1 Score") %>%
  bg(part = "body", bg = "lightgray", i = ~ Metric == "AUC") %>%
  align(j = 2:3, align = "center", part = "body") %>%
  autofit()

# Print the flextable
comparison_table

# ---------------------- Bar Graph Comparison ----------------------

# Reshape the data for plotting
metrics_long <- reshape2::melt(metrics_comparison, id.vars = "Metric", variable.name = "Model", value.name = "Value")

# Create the bar graph
ggplot(metrics_long, aes(x = Metric, y = Value, fill = Model)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("Logistic_Regression" = "blue", "Random_Forest" = "red")) +
  labs(title = "Model Comparison - Logistic Regression vs Random Forest", 
       x = "Metric", y = "Value", fill = "Model") +
  theme_minimal()

```
### Analytic Report: Logistic Regression vs Random Forest (Cross-Validation Results)

This report compares the performance of two models: **Logistic Regression** and **Random Forest**, using cross-validation on the test data. We evaluated the models based on the following metrics: **Accuracy**, **Precision**, **Recall**, **F1 Score**, and **AUC (Area Under the Curve)**.

#### Overall Findings:

- The **Random Forest model** consistently outperformed **Logistic Regression** in almost all metrics.
- **Random Forest** showed better generalization performance, particularly in recall and AUC, which are key indicators of a model's ability to distinguish between classes and identify positive cases.
- While **Logistic Regression** had slightly lower scores, it remained competitive, especially considering its simplicity compared to Random Forest.

#### Metric Comparisons:

1. **Accuracy**:
   - Logistic Regression: **0.7972**
   - Random Forest: **0.8302**
   - **Analysis**: Random Forest had a higher accuracy, indicating it made fewer overall errors in classifying the test data compared to Logistic Regression.

2. **Precision**:
   - Logistic Regression: **0.7384**
   - Random Forest: **0.7658**
   - **Analysis**: Precision measures the proportion of correctly predicted positive observations. Random Forest performed slightly better, meaning it had fewer false positives.

3. **Recall**:
   - Logistic Regression: **0.8208**
   - Random Forest: **0.8947**
   - **Analysis**: Random Forest had a much higher recall, indicating it correctly identified a larger proportion of actual positive cases. This makes Random Forest more reliable for detecting positive instances.

4. **F1 Score**:
   - Logistic Regression: **0.8018**
   - Random Forest: **0.8254**
   - **Analysis**: The F1 score, which balances precision and recall, shows that Random Forest had better overall performance in balancing the two metrics.

5. **AUC (Area Under the Curve)**:
   - Logistic Regression: **0.8907**
   - Random Forest: **0.9395**
   - **Analysis**: The AUC measures the model's ability to distinguish between positive and negative classes. A higher AUC means better discriminatory power. Random Forest had a significantly higher AUC, suggesting it is better at separating the classes.

#### Key Observations:

- **Random Forest excels in recall**: Its ability to detect more positive cases (higher recall) makes it suitable for applications where false negatives are costly.
- **Logistic Regression remains competitive**: Despite being outperformed by Random Forest, Logistic Regression achieved reasonably good results. Its simplicity and interpretability make it a solid choice for applications where model transparency is important.
- **Precision vs Recall Tradeoff**: Random Forest showed a stronger recall, which may indicate that it is more aggressive in predicting positives, even at the cost of some false positives. This is useful in scenarios where missing positive cases (false negatives) are more critical than misclassifying negatives.

#### Conclusion:
- **Random Forest** is the preferred model based on its superior performance across all metrics, particularly in recall and AUC.
- **Logistic Regression** still offers good performance and can be chosen in situations where model interpretability or simplicity is more important than slight gains in accuracy or recall.