### 2024-12-03 R Projects/CAPM CASE STUDY 3.qmd

---
title: "Case Study Down Jones"
author: "Leonel Salazar, Seth Harris, Joaquin Ramirez, Collin Real"
format: docx
---

```{r echo=FALSE}

library(tidyverse)
library(here)
library(ggplot2)
library(gridExtra)
library(tseries)
library(dplyr)
library(quantmod)

here()

```

```{r echo=FALSE, warning=FALSE, error=FALSE}

# Use the here function to construct the file path and import the dataset
data <- read.csv(here("dow_jones_index.data.csv"), header = TRUE, sep = ",")

head(data)

str(data)

```

```{r echo=FALSE, warning=FALSE, error=FALSE}

# Check for missing data or NA values in the dataset

missing_data_summary <- colSums(is.na(data))
missing_data_summary
```

```{r echo=FALSE, warning=FALSE, error=FALSE}

# removed.

data_clean <- data

```

```{r echo=FALSE, warning=FALSE, error=FALSE}

# Convert 'quarter' variable to a factor
data_clean$quarter <- as.factor(data_clean$quarter)

# Convert 'stock' variable to a factor
data_clean$stock <- as.factor(data_clean$stock)

# Convert 'date' variable to Date class format
data_clean$date <- as.Date(data_clean$date, format = "%m/%d/%Y")

# Remove non-numeric characters from 'open', 'high', 'low', and 'close' before converting to numeric
data_clean$open <- as.numeric(gsub("[^0-9.]", "", data_clean$open))
data_clean$high <- as.numeric(gsub("[^0-9.]", "", data_clean$high))
data_clean$low <- as.numeric(gsub("[^0-9.]", "", data_clean$low))
data_clean$close <- as.numeric(gsub("[^0-9.]", "", data_clean$close))

# Convert 'volume' variable to numeric
data_clean$volume <- as.numeric(gsub("[^0-9.]", "", data_clean$volume))

# # Convert 'percent_change_price' to numeric by removing '%' and dividing by 100
# data_clean$percent_change_price <- as.numeric(gsub("%", "", data_clean$percent_change_price)) / 100
# 
# Convert 'percent_change_volume_over_last_wk' to numeric by removing '%'
data_clean$percent_change_volume_over_last_wk <- as.numeric(gsub("%", "", data_clean$percent_change_volume_over_last_wk))

# Remove non-numeric characters from 'previous_weeks_volume', 'next_weeks_open', and 'next_weeks_close' before converting to numeric
data_clean$previous_weeks_volume <- as.numeric(gsub("[^0-9.]", "", data_clean$previous_weeks_volume))
data_clean$next_weeks_open <- as.numeric(gsub("[^0-9.]", "", data_clean$next_weeks_open))
data_clean$next_weeks_close <- as.numeric(gsub("[^0-9.]", "", data_clean$next_weeks_close))

# # Convert 'percent_change_next_weeks_price' to numeric by removing '%' and dividing by 100
# data_clean$percent_change_next_weeks_price <- as.numeric(gsub("%", "", data_clean$percent_change_next_weeks_price)) / 100

# Convert 'days_to_next_dividend' to numeric
data_clean$days_to_next_dividend <- as.numeric(gsub("[^0-9.]", "", data_clean$days_to_next_dividend))

str(data_clean)

```

```{r echo=FALSE, warning=FALSE, error=FALSE}
# Step 1: Inspect the unique stock symbols
unique_stocks <- unique(data_clean$stock) # Adjust column name if different

# Step 2: Randomly select 5 unique stocks
set.seed(123) # Set seed for reproducibility
selected_stocks <- sample(unique_stocks, 30)

# Step 3: Create a directory for each stock and save its data
for (stock in selected_stocks) {
  # Filter data for the current stock
  stock_data <- data_clean %>%
    filter(stock == !!stock) %>%
    select(stock,date, open, high, low, close, volume) # Adjust column names to match your dataset

   # Create file path for each stock
  file_name <- here( paste0("stock_", stock, ".csv"))

  # Save the stock data as a CSV file within its folder
  file_name <- here( paste0("stock_", stock, ".csv"))
  write.csv(stock_data, file_name, row.names = FALSE)
}
```

```{r echo=FALSE, warning=FALSE, error=FALSE}
# Step 1: Extract unique stock names from the main dataset
stock_symbols <- unique(data_clean$stock) # Ensure 'data_clean' is already processed

# Step 2: Load data for each stock into individual variables
for (symbol in stock_symbols) {
  # Construct file name dynamically
  file_path <- here(paste0("stock_", symbol, ".csv"))

  # Check if the file exists
  if (file.exists(file_path)) {
    # Read the stock data
    assign(symbol, read.csv(file_path, header = TRUE, sep = ","))
    cat("Loaded data for stock:", symbol, "\n")
  } else {
    warning(paste("File not found for stock:", symbol))
  }
}

# Example: Check the structure of one dataset (e.g., AAPL)
if (exists("AAPL")) {
  str(AAPL)
}

```

```{r echo=FALSE, warning=FALSE, error=FALSE}

# Step 2: Select only the relevant variables
clean_stock <- data_clean %>%
  select(stock,date, open, high, low, close, volume) # Adjust column names to match your dataset

# Step 3: Save the cleaned dataset as dowjones.csv
output_path <- here("clean_stock.csv")
write.csv(data_clean, output_path, row.names = FALSE)

clean_stock <- read.csv(here("clean_stock.csv"), header = TRUE, sep = ",")

str(clean_stock)
```

```{r echo=FALSE, warning=FALSE, error=FALSE}

str(GE)

str(clean_stock)

```

```{r echo=FALSE, warning=FALSE, error=FALSE}

# Convert 'date' column to Date format and numeric conversion for the other columns
clean_stock <- data %>%
  mutate(
    date = as.Date(date, format = "%m/%d/%Y"), # Convert 'date' to Date format
    open = as.numeric(gsub("\\$", "", open)),  # Remove dollar sign and convert to numeric
    high = as.numeric(gsub("\\$", "", high)),  # Remove dollar sign and convert to numeric
    low = as.numeric(gsub("\\$", "", low)),    # Remove dollar sign and convert to numeric
    close = as.numeric(gsub("\\$", "", close)),# Remove dollar sign and convert to numeric
    volume = as.numeric(volume)                # Ensure 'volume' is numeric
  )

# Remove the 'stock' column
clean_stock <- clean_stock %>%
  select(date, open, high, low, close, volume)

str(clean_stock)
```

```{r}

library(dplyr)

# Ensure columns are properly formatted
clean_stock <- clean_stock %>%
  mutate(
    date = as.Date(date, format = "%m/%d/%Y"),
    open = as.numeric(gsub("\\$", "", open)),
    high = as.numeric(gsub("\\$", "", high)),
    low = as.numeric(gsub("\\$", "", low)),
    close = as.numeric(gsub("\\$", "", close)),
    volume = as.numeric(volume)
  )

# Check for issues in data conversion
print(summary(data))  # Provides an overview of the data

# Group by 'date' and calculate averages, handling NAs
averaged_data <- clean_stock %>%
  group_by(date) %>%
  summarize(
    avg_open = mean(open, na.rm = TRUE),
    avg_high = mean(high, na.rm = TRUE),
    avg_low = mean(low, na.rm = TRUE),
    avg_close = mean(close, na.rm = TRUE),
    avg_volume = mean(volume, na.rm = TRUE)
  )

# View the cleaned data
print(head(averaged_data))
```
```{r echo=FALSE, warning=FALSE, error=FALSE}

# Step 1: Define stock symbols
stock_symbols <- unique(data_clean$stock)

# Step 2: Remove the 'stock' column for each dataset dynamically
for (symbol in stock_symbols) {
  # Check if the dataset exists in the environment
  if (exists(symbol)) {
    # Get the dataset
    stock_data <- get(symbol)

    # Remove the 'stock' column
    stock_data <- stock_data %>% select(-stock)

    # Save the modified dataset back to its original variable
    assign(symbol, stock_data)

    # Print confirmation
    cat("Removed 'stock' column for:", symbol, "\n")
  } else {
    warning(paste("Dataset not found for stock:", symbol))
  }
}

```

```{r echo=FALSE, warning=FALSE, error=FALSE}
here()

# Load the data files for all stocks
stock_AA <- read.csv(here("Week 6", "stock_AA.csv"), stringsAsFactors = FALSE)
stock_AXP <- read.csv(here("Week 6", "stock_AXP.csv"), stringsAsFactors = FALSE)
stock_BA <- read.csv(here("Week 6", "stock_BA.csv"), stringsAsFactors = FALSE)
stock_BAC <- read.csv(here("Week 6", "stock_BAC.csv"), stringsAsFactors = FALSE)
stock_CAT <- read.csv(here("Week 6", "stock_CAT.csv"), stringsAsFactors = FALSE)
stock_CSCO <- read.csv(here("Week 6", "stock_CSCO.csv"), stringsAsFactors = FALSE)
stock_CVX <- read.csv(here("Week 6", "stock_CVX.csv"), stringsAsFactors = FALSE)
stock_DIS <- read.csv(here("Week 6", "stock_DIS.csv"), stringsAsFactors = FALSE)
stock_DD <- read.csv(here("Week 6", "stock_DD.csv"), stringsAsFactors = FALSE)
stock_GE <- read.csv(here("Week 6", "stock_GE.csv"), stringsAsFactors = FALSE)
stock_HD <- read.csv(here("Week 6", "stock_HD.csv"), stringsAsFactors = FALSE)
stock_HPQ <- read.csv(here("Week 6", "stock_HPQ.csv"), stringsAsFactors = FALSE)
stock_IBM <- read.csv(here("Week 6", "stock_IBM.csv"), stringsAsFactors = FALSE)
stock_INTC <- read.csv(here("Week 6", "stock_INTC.csv"), stringsAsFactors = FALSE)
stock_JNJ <- read.csv(here("Week 6", "stock_JNJ.csv"), stringsAsFactors = FALSE)
stock_JPM <- read.csv(here("Week 6", "stock_JPM.csv"), stringsAsFactors = FALSE)
stock_KO <- read.csv(here("Week 6", "stock_KO.csv"), stringsAsFactors = FALSE)
stock_KRFT <- read.csv(here("Week 6", "stock_KRFT.csv"), stringsAsFactors = FALSE)
stock_MCD <- read.csv(here("Week 6", "stock_MCD.csv"), stringsAsFactors = FALSE)
stock_MMM <- read.csv(here("Week 6", "stock_MMM.csv"), stringsAsFactors = FALSE)
stock_MRK <- read.csv(here("Week 6", "stock_MRK.csv"), stringsAsFactors = FALSE)
stock_MSFT <- read.csv(here("Week 6", "stock_MSFT.csv"), stringsAsFactors = FALSE)
stock_PFE <- read.csv(here("Week 6", "stock_PFE.csv"), stringsAsFactors = FALSE)
stock_PG <- read.csv(here("Week 6", "stock_PG.csv"), stringsAsFactors = FALSE)
stock_TRV <- read.csv(here("Week 6", "stock_TRV.csv"), stringsAsFactors = FALSE)
stock_UTX <- read.csv(here("Week 6", "stock_UTX.csv"), stringsAsFactors = FALSE)
stock_VZ <- read.csv(here("Week 6", "stock_VZ.csv"), stringsAsFactors = FALSE)
stock_WMT <- read.csv(here("Week 6", "stock_WMT.csv"), stringsAsFactors = FALSE)
stock_XOM <- read.csv(here("Week 6", "stock_XOM.csv"), stringsAsFactors = FALSE)

# Save the cleaned data back to their respective files
write.csv(stock_AA, here("Week 6", "stock_AA.csv"), row.names = FALSE)
write.csv(stock_AXP, here("Week 6", "stock_AXP.csv"), row.names = FALSE)
write.csv(stock_BA, here("Week 6", "stock_BA.csv"), row.names = FALSE)
write.csv(stock_BAC, here("Week 6", "stock_BAC.csv"), row.names = FALSE)
write.csv(stock_CAT, here("Week 6", "stock_CAT.csv"), row.names = FALSE)
write.csv(stock_CSCO, here("Week 6", "stock_CSCO.csv"), row.names = FALSE)
write.csv(stock_CVX, here("Week 6", "stock_CVX.csv"), row.names = FALSE)
write.csv(stock_DIS, here("Week 6", "stock_DIS.csv"), row.names = FALSE)
write.csv(stock_DD, here("Week 6", "stock_DD.csv"), row.names = FALSE)
write.csv(stock_GE, here("Week 6", "stock_GE.csv"), row.names = FALSE)
write.csv(stock_HD, here("Week 6", "stock_HD.csv"), row.names = FALSE)
write.csv(stock_HPQ, here("Week 6", "stock_HPQ.csv"), row.names = FALSE)
write.csv(stock_IBM, here("Week 6", "stock_IBM.csv"), row.names = FALSE)
write.csv(stock_INTC, here("Week 6", "stock_INTC.csv"), row.names = FALSE)
write.csv(stock_JNJ, here("Week 6", "stock_JNJ.csv"), row.names = FALSE)
write.csv(stock_JPM, here("Week 6", "stock_JPM.csv"), row.names = FALSE)
write.csv(stock_KO, here("Week 6", "stock_KO.csv"), row.names = FALSE)
write.csv(stock_KRFT, here("Week 6", "stock_KRFT.csv"), row.names = FALSE)
write.csv(stock_MCD, here("Week 6", "stock_MCD.csv"), row.names = FALSE)
write.csv(stock_MMM, here("Week 6", "stock_MMM.csv"), row.names = FALSE)
write.csv(stock_MRK, here("Week 6", "stock_MRK.csv"), row.names = FALSE)
write.csv(stock_MSFT, here("Week 6", "stock_MSFT.csv"), row.names = FALSE)
write.csv(stock_PFE, here("Week 6", "stock_PFE.csv"), row.names = FALSE)
write.csv(stock_PG, here("Week 6", "stock_PG.csv"), row.names = FALSE)
write.csv(stock_TRV, here("Week 6", "stock_TRV.csv"), row.names = FALSE)
write.csv(stock_UTX, here("Week 6", "stock_UTX.csv"), row.names = FALSE)
write.csv(stock_VZ, here("Week 6", "stock_VZ.csv"), row.names = FALSE)
write.csv(stock_WMT, here("Week 6", "stock_WMT.csv"), row.names = FALSE)
write.csv(stock_XOM, here("Week 6", "stock_XOM.csv"), row.names = FALSE)

cat("Data cleaning completed for all stocks.\n")

```

```{r echo=FALSE, warning=FALSE, error=FALSE}

# Remove the 'stock' column for all stocks
stock_AA <- stock_AA %>% select(-stock)
stock_AXP <- stock_AXP %>% select(-stock)
stock_BA <- stock_BA %>% select(-stock)
stock_BAC <- stock_BAC %>% select(-stock)
stock_CAT <- stock_CAT %>% select(-stock)
stock_CSCO <- stock_CSCO %>% select(-stock)
stock_CVX <- stock_CVX %>% select(-stock)
stock_DIS <- stock_DIS %>% select(-stock)
stock_DD <- stock_DD %>% select(-stock)
stock_GE <- stock_GE %>% select(-stock)
stock_HD <- stock_HD %>% select(-stock)
stock_HPQ <- stock_HPQ %>% select(-stock)
stock_IBM <- stock_IBM %>% select(-stock)
stock_INTC <- stock_INTC %>% select(-stock)
stock_JNJ <- stock_JNJ %>% select(-stock)
stock_JPM <- stock_JPM %>% select(-stock)
stock_KO <- stock_KO %>% select(-stock)
stock_KRFT <- stock_KRFT %>% select(-stock)
stock_MCD <- stock_MCD %>% select(-stock)
stock_MMM <- stock_MMM %>% select(-stock)
stock_MRK <- stock_MRK %>% select(-stock)
stock_MSFT <- stock_MSFT %>% select(-stock)
stock_PFE <- stock_PFE %>% select(-stock)
stock_PG <- stock_PG %>% select(-stock)
stock_TRV <- stock_TRV %>% select(-stock)
stock_UTX <- stock_UTX %>% select(-stock)
stock_VZ <- stock_VZ %>% select(-stock)
stock_WMT <- stock_WMT %>% select(-stock)
stock_XOM <- stock_XOM %>% select(-stock)

cat("Removed the 'stock' column for all stocks.\n")

```

```{r echo=FALSE, warning=FALSE, error=FALSE}

```

```{r echo=FALSE, warning=FALSE, error=FALSE}

# Compute returns and remove missing values for all stocks
ReturnAA <- na.omit(Delt(stock_AA$close))
ReturnAXP <- na.omit(Delt(stock_AXP$close))
ReturnBA <- na.omit(Delt(stock_BA$close))
ReturnBAC <- na.omit(Delt(stock_BAC$close))
ReturnCAT <- na.omit(Delt(stock_CAT$close))
ReturnCSCO <- na.omit(Delt(stock_CSCO$close))
ReturnCVX <- na.omit(Delt(stock_CVX$close))
ReturnDIS <- na.omit(Delt(stock_DIS$close))
ReturnDD <- na.omit(Delt(stock_DD$close))
ReturnGE <- na.omit(Delt(stock_GE$close))
ReturnHD <- na.omit(Delt(stock_HD$close))
ReturnHPQ <- na.omit(Delt(stock_HPQ$close))
ReturnIBM <- na.omit(Delt(stock_IBM$close))
ReturnINTC <- na.omit(Delt(stock_INTC$close))
ReturnJNJ <- na.omit(Delt(stock_JNJ$close))
ReturnJPM <- na.omit(Delt(stock_JPM$close))
ReturnKO <- na.omit(Delt(stock_KO$close))
ReturnKRFT <- na.omit(Delt(stock_KRFT$close))
ReturnMCD <- na.omit(Delt(stock_MCD$close))
ReturnMMM <- na.omit(Delt(stock_MMM$close))
ReturnMRK <- na.omit(Delt(stock_MRK$close))
ReturnMSFT <- na.omit(Delt(stock_MSFT$close))
ReturnPFE <- na.omit(Delt(stock_PFE$close))
ReturnPG <- na.omit(Delt(stock_PG$close))
ReturnTRV <- na.omit(Delt(stock_TRV$close))
ReturnUTX <- na.omit(Delt(stock_UTX$close))
ReturnVZ <- na.omit(Delt(stock_VZ$close))
ReturnWMT <- na.omit(Delt(stock_WMT$close))
ReturnXOM <- na.omit(Delt(stock_XOM$close))

# Compute returns for averaged data (Dow Jones index)
ReturnStock <- na.omit(Delt(averaged_data$avg_close))

cat("Computed returns and removed missing values for all stocks.\n")

```

```{r echo=FALSE, warning=FALSE, error=FALSE}

# Merge the returns of all three index into single data
MyData = cbind(ReturnBA, ReturnGE, ReturnINTC, ReturnJNJ, ReturnMCD, ReturnStock)
```

```{r echo=FALSE, warning=FALSE, error=FALSE}

# Merge the returns of all stocks into a single data frame
MyData <- cbind(
  ReturnAA, ReturnAXP, ReturnBA, ReturnBAC, ReturnCAT, 
  ReturnCSCO, ReturnCVX, ReturnDIS, ReturnDD, ReturnGE,
  ReturnHD, ReturnHPQ, ReturnIBM, ReturnINTC, ReturnJNJ,
  ReturnJPM, ReturnKO, ReturnKRFT, ReturnMCD, ReturnMMM,
  ReturnMRK, ReturnMSFT, ReturnPFE, ReturnPG, ReturnTRV,
  ReturnUTX, ReturnVZ, ReturnWMT, ReturnXOM, ReturnStock
)

# Set column names, excluding NKE
colnames(MyData) <- c(
  "AA", "AXP", "BA", "BAC", "CAT", 
  "CSCO", "CVX", "DIS", "DD", "GE",
  "HD", "HPQ", "IBM", "INTC", "JNJ",
  "JPM", "KO", "KRFT", "MCD", "MMM",
  "MRK", "MSFT", "PFE", "PG", "TRV",
  "UTX", "VZ", "WMT", "XOM", "Dow"
)

# Display the first few rows of the combined data
head(MyData)

```

```{r echo=FALSE, warning=FALSE, error=FALSE}

# See how the data looks. You can see the risk is much lesser in the S&P as this 
# represents the market, compared to the individual stock. 
boxplot(MyData,main="Expected Return", xlab="Stock Picks", ylab="Return")

```
```{r echo=FALSE, warning=FALSE, error=FALSE}

# Number of stocks to compare in each plot
stocks_per_plot <- 4

# Get the column names excluding "Dow"
stock_columns <- colnames(MyData)[colnames(MyData) != "Dow"]

# Loop through the stocks in chunks of 4 and create boxplots
for (i in seq(1, length(stock_columns), by = stocks_per_plot)) {
  # Subset the stocks for the current plot and include Dow
  selected_columns <- c(stock_columns[i:min(i + stocks_per_plot - 1, length(stock_columns))], "Dow")

  # Create a boxplot for the selected stocks
  boxplot(
    MyData[, selected_columns, drop = FALSE],
    main = paste("Expected Return for Stocks", paste(selected_columns, collapse = ", ")),
    xlab = "Stock Picks",
    ylab = "Return"
  )
}

```
```{r}

# Load the here package
library(here)

# Define the number of stocks to compare in each plot
stocks_per_plot <- 4

# Get the column names excluding "Dow"
stock_columns <- colnames(MyData)[colnames(MyData) != "Dow"]

# Define the output folder using the here package
output_folder <- here("plots")

# Create the folder if it doesn't exist
if (!dir.exists(output_folder)) {
  dir.create(output_folder, recursive = TRUE)
}

# Loop through the stocks in chunks of 4 and create boxplots
for (i in seq(1, length(stock_columns), by = stocks_per_plot)) {
  # Subset the stocks for the current plot and include Dow
  selected_columns <- c(stock_columns[i:min(i + stocks_per_plot - 1, length(stock_columns))], "Dow")

  # Define the filename for the current plot
  plot_filename <- file.path(output_folder, paste0("Boxplot_", i, ".png"))

  # Save the boxplot as a PNG file
  png(plot_filename, width = 800, height = 600)
  boxplot(
    MyData[, selected_columns, drop = FALSE],
    main = paste("Expected Return for Stocks", paste(selected_columns, collapse = ", ")),
    xlab = "Stock Picks",
    ylab = "Return"
  )
  dev.off()  # Close the graphics device
}

```

```{r echo=FALSE, warning=FALSE, error=FALSE}

# Compute mean and stdev for the returns.
DataMean=apply(MyData, 2, mean)
DataSD=apply(MyData, 2, sd)
# Take a look at the means and standard deviations. 
cbind(DataMean,DataSD)
```
```{r echo=FALSE, warning=FALSE, error=FALSE}

# List of stocks to process
stocks <- colnames(MyData)[colnames(MyData) != "Dow"] # Exclude Dow

# Initialize an empty data frame to store results
results <- data.frame(Stock = character(), Beta = numeric(), R_Squared = numeric(), stringsAsFactors = FALSE)

# Loop through each stock and compute the linear model
for (stock in stocks) {
  # Construct the formula dynamically
  formula <- as.formula(paste(stock, "~ Dow"))

  # Fit the linear model
  lm_model <- lm(formula, data = as.data.frame(MyData))

  # Extract Beta (coefficient for Dow) and R-squared
  beta <- summary(lm_model)$coefficients[2, 1]
  r_squared <- summary(lm_model)$r.squared

  # Append results to the data frame
  results <- rbind(results, data.frame(Stock = stock, Beta = beta, R_Squared = r_squared))
}

# Order results by Beta (safest to riskiest)
results <- results[order(results$Beta), ]

# Print as a long list
for (i in 1:nrow(results)) {
  cat(paste0(
    "Stock: ", results$Stock[i], 
    ", Beta: ", round(results$Beta[i], 4), 
    ", R-Squared: ", round(results$R_Squared[i], 4), "\n"
  ))
}

# Save the ordered results table to a CSV file (optional)
write.csv(results, file = here("Week 6", "stock_beta_results_sorted.csv"), row.names = FALSE)

```
```{r echo=FALSE, warning=FALSE, error=FALSE}

# Load the required library
library(flextable)

# List of stocks to process
stocks <- colnames(MyData)[colnames(MyData) != "Dow"] # Exclude Dow

# Initialize an empty data frame to store results
results <- data.frame(Stock = character(), Beta = numeric(), R_Squared = numeric(), Dow_Beta = numeric(), stringsAsFactors = FALSE)

# Loop through each stock and compute the linear model
for (stock in stocks) {
  # Construct the formula dynamically
  formula <- as.formula(paste(stock, "~ Dow"))

  # Fit the linear model
  lm_model <- lm(formula, data = as.data.frame(MyData))

  # Extract Beta (coefficient for Dow) and R-squared
  beta <- summary(lm_model)$coefficients[2, 1]
  r_squared <- summary(lm_model)$r.squared

  # Append results to the data frame
  results <- rbind(results, data.frame(Stock = stock, Beta = beta, R_Squared = r_squared, Dow_Beta = 1)) # Dow_Beta is always 1
}

# Order results by Beta (safest to riskiest)
results <- results[order(results$Beta), ]

# Create the flextable
ft <- flextable(results) %>%
  # Add a title
  set_caption("Beta vs Dow Beta") %>%
  # Highlight the Dow Beta column in light blue
  bg(j = "Dow_Beta", bg = "lightblue", part = "body") %>%
  # Format numeric values
  colformat_double(j = c("Beta", "R_Squared", "Dow_Beta"), digits = 4) %>%
  # Autofit column sizes
  autofit()

# Print the table
ft

```

```{r}

# Load the required libraries
library(flextable)
library(here)
library(webshot)

# Install webshot dependencies if not already installed
# webshot::install_phantomjs()

# Define the output folder using the here package
output_folder <- here("tables")

# Create the folder if it doesn't exist
if (!dir.exists(output_folder)) {
  dir.create(output_folder, recursive = TRUE)
}

# List of stocks to process (excluding "Dow")
stocks <- colnames(MyData)[colnames(MyData) != "Dow"]

# Initialize an empty data frame to store results
results <- data.frame(Stock = character(), Beta = numeric(), R_Squared = numeric(), Dow_Beta = numeric(), stringsAsFactors = FALSE)

# Loop through each stock and compute the linear model
for (stock in stocks) {
  # Construct the formula dynamically
  formula <- as.formula(paste(stock, "~ Dow"))

  # Fit the linear model
  lm_model <- lm(formula, data = as.data.frame(MyData))

  # Extract Beta (coefficient for Dow) and R-squared
  beta <- summary(lm_model)$coefficients[2, 1]
  r_squared <- summary(lm_model)$r.squared

  # Append results to the data frame
  results <- rbind(results, data.frame(Stock = stock, Beta = beta, R_Squared = r_squared, Dow_Beta = 1)) # Dow_Beta is always 1
}

# Order results by Beta (safest to riskiest)
results <- results[order(results$Beta), ]

# Split results into chunks of stocks per table
stocks_per_table <- 10
num_tables <- ceiling(nrow(results) / stocks_per_table)

# Loop to create and save each table as an image
for (i in seq(1, num_tables)) {
  # Subset the data for the current table
  start_row <- (i - 1) * stocks_per_table + 1
  end_row <- min(i * stocks_per_table, nrow(results))
  table_data <- results[start_row:end_row, ]

  # Create the flextable for the current chunk
  ft <- flextable(table_data) %>%
    set_caption(paste("Beta vs Dow Beta (Table", i, ")")) %>%
    bg(j = "Dow_Beta", bg = "lightblue", part = "body") %>%
    colformat_double(j = c("Beta", "R_Squared", "Dow_Beta"), digits = 4) %>%
    autofit()

  # Define the filename for the current table
  image_filename <- file.path(output_folder, paste0("Table_", i, ".png"))

  # Save the flextable as an image
  save_as_image(ft, path = image_filename)
}

```

```{r echo=FALSE, warning=FALSE, error=FALSE}

# Load the required library
library(flextable)

# Inputs for CAPM
risk_free_rate <- 0.03 # Example: 3% annualized (e.g., 10-year Treasury bond yield)
market_return <- 0.08  # Example: 8% annualized expected return of Dow Jones index
market_risk_premium <- market_return - risk_free_rate

# Assuming `results` contains Beta values calculated earlier
# Add Expected Return based on the CAPM formula
results$Expected_Return <- risk_free_rate + (results$Beta * market_risk_premium)

# Order by Expected Return (safest to riskiest stock based on CAPM)
results <- results[order(results$Expected_Return), ]

# Create a flextable for results
ft <- flextable(results) %>%
  # Add a title
  set_caption("CAPM: Expected Returns vs Dow Beta") %>%
  # Highlight the Dow Beta column in light blue
  bg(j = "Dow_Beta", bg = "lightblue", part = "body") %>%
  # Format numeric columns
  colformat_double(j = c("Beta", "R_Squared", "Dow_Beta", "Expected_Return"), digits = 4) %>%
  # Autofit column sizes
  autofit()

# Display the flextable
ft

# Save the results table to a CSV file (optional)
write.csv(results, file = here("Week 6", "stock_capm_results.csv"), row.names = FALSE)

# Print a message to indicate successful calculation
cat("CAPM calculations and table generation completed successfully.\n")

```

```{r}

```

```{r echo=FALSE, warning=FALSE, error=FALSE}
# Ensure MyData is a data frame
MyData <- as.data.frame(MyData)

# Load required libraries
library(ggplot2)
library(patchwork) # For arranging plots in a grid

# Exclude "Dow" from the stock list
stock_columns <- colnames(MyData)[colnames(MyData) != "Dow"]

# Initialize an empty list to store plots
plot_list <- list()

# Loop through each stock to generate plots
for (stock in stock_columns) {
  # Create a data frame for the current stock and the Dow
  plot_data <- data.frame(
    Market_Excess_Return = MyData$Dow, # Dow is the market
    Stock_Excess_Return = MyData[[stock]]
  )

  # Generate the scatter plot with regression line
  p <- ggplot(plot_data, aes(x = Market_Excess_Return, y = Stock_Excess_Return)) +
    geom_point(color = "blue", alpha = 0.6) +  # Scatter plot points
    geom_smooth(method = "lm", color = "red", se = FALSE) +  # Regression line
    labs(
      title = paste(stock, "vs. Dow Jones"),
      x = "Weekly Market Excess Return (%)",
      y = "Weekly Stock Excess Return (%)"
    ) +
    theme_minimal() +
    annotate(
      "text",
      x = max(plot_data$Market_Excess_Return, na.rm = TRUE) * 0.7,
      y = max(plot_data$Stock_Excess_Return, na.rm = TRUE) * 0.7,
      label = paste(
        "Alpha =", round(coef(lm(Stock_Excess_Return ~ Market_Excess_Return, data = plot_data))[1], 4),
        "\nBeta =", round(coef(lm(Stock_Excess_Return ~ Market_Excess_Return, data = plot_data))[2], 4)
      ),
      color = "black"
    )

  # Append the plot to the list
  plot_list[[stock]] <- p
}

# Combine plots into a grid in batches of 3
for (i in seq(1, length(plot_list), by = 3)) {
  # Select up to 3 plots for the current batch
  selected_plots <- plot_list[i:min(i + 2, length(plot_list))]

  # Combine the selected plots into a 1 x 3 grid
  combined_plot <- wrap_plots(selected_plots, ncol = 3)

  # Save the combined plot as a PNG
  ggsave(
    filename = paste0("Week_6_Batch_", ceiling(i / 3), "_Graphs.png"),
    plot = combined_plot,
    width = 12, height = 4
  )
}

```
```{r}

# Load required libraries
library(flextable)
library(here)
library(webshot)

# Install webshot dependencies if not already installed
# webshot::install_phantomjs()

# Inputs for CAPM
risk_free_rate <- 0.03  # Example: 3% annualized (e.g., 10-year Treasury bond yield)
market_return <- 0.08   # Example: 8% annualized expected return of Dow Jones index
market_risk_premium <- market_return - risk_free_rate

# Assuming `results` contains Beta values calculated earlier
# Add Expected Return based on the CAPM formula
results$Expected_Return <- risk_free_rate + (results$Beta * market_risk_premium)

# Order by Expected Return (safest to riskiest stock based on CAPM)
results <- results[order(results$Expected_Return), ]

# Create a flextable for results with proper styling
ft <- flextable(results) %>%
  set_caption("CAPM: Expected Returns vs Dow Beta") %>%
  bg(bg = "white", part = "all") %>%  # Set white background for the entire table
  color(color = "black", part = "all") %>%  # Set text color to black
  border_outer() %>%
  colformat_double(j = c("Beta", "R_Squared", "Dow_Beta", "Expected_Return"), digits = 4) %>%
  autofit()

# Define the output folder for the image
output_folder <- here("Week 6")
if (!dir.exists(output_folder)) {
  dir.create(output_folder, recursive = TRUE)
}

# Define the image path
image_path <- file.path(output_folder, "CAPM_Results_Table.png")

# Save the flextable as an image with proper rendering
save_as_image(ft, path = image_path)

# Print a message to indicate successful saving
```

```{r echo=FALSE, warning=FALSE, error=FALSE}

# Stock Analysis Report

# This report includes visualizations for the analysis of stock performance data. Below are the charts for each batch of stocks analyzed.

## Visualizations

# Load knitr package
library(knitr)

# Display all images in a single code chunk
knitr::include_graphics(c(
  "Week_6_Batch_1_Graphs.png",
  "Week_6_Batch_2_Graphs.png",
  "Week_6_Batch_3_Graphs.png",
  "Week_6_Batch_4_Graphs.png",
  "Week_6_Batch_5_Graphs.png",
  "Week_6_Batch_6_Graphs.png",
  "Week_6_Batch_7_Graphs.png",
  "Week_6_Batch_8_Graphs.png",
  "Week_6_Batch_9_Graphs.png",
  "Week_6_Batch_10_Graphs.png"
))

```