### 2024-12-03 R Projects/CAPM_CASESTUDY.qmd

---
title: "Case Study Down Jones"
author: "Leonel Salazar, Seth Harris, Joaquin Ramirez, Collin Real"
format: docx
---

```{r}

library(tidyverse)
library(here)
library(ggplot2)
library(gridExtra)
library(tseries)
library(dplyr)
library(quantmod)

```

```{r}

# Use the here function to construct the file path and import the dataset
data <- read.csv(here("dow_jones_index.data.csv"), header = TRUE, sep = ",")

head(data)

str(data)

```

```{r}

# Check for missing data or NA values in the dataset

missing_data_summary <- colSums(is.na(data))
missing_data_summary
```

```{r}

# Remove rows with missing data and create a new dataset named data_clean
data_clean <- na.omit(data)

str(data_clean)
```

```{r}

# Convert 'quarter' variable to a factor
data_clean$quarter <- as.factor(data_clean$quarter)

# Convert 'stock' variable to a factor
data_clean$stock <- as.factor(data_clean$stock)

# Convert 'date' variable to Date class format
data_clean$date <- as.Date(data_clean$date, format = "%m/%d/%Y")

# Remove non-numeric characters from 'open', 'high', 'low', and 'close' before converting to numeric
data_clean$open <- as.numeric(gsub("[^0-9.]", "", data_clean$open))
data_clean$high <- as.numeric(gsub("[^0-9.]", "", data_clean$high))
data_clean$low <- as.numeric(gsub("[^0-9.]", "", data_clean$low))
data_clean$close <- as.numeric(gsub("[^0-9.]", "", data_clean$close))

# Convert 'volume' variable to numeric
data_clean$volume <- as.numeric(gsub("[^0-9.]", "", data_clean$volume))

# # Convert 'percent_change_price' to numeric by removing '%' and dividing by 100
# data_clean$percent_change_price <- as.numeric(gsub("%", "", data_clean$percent_change_price)) / 100
# 
# Convert 'percent_change_volume_over_last_wk' to numeric by removing '%'
data_clean$percent_change_volume_over_last_wk <- as.numeric(gsub("%", "", data_clean$percent_change_volume_over_last_wk))

# Remove non-numeric characters from 'previous_weeks_volume', 'next_weeks_open', and 'next_weeks_close' before converting to numeric
data_clean$previous_weeks_volume <- as.numeric(gsub("[^0-9.]", "", data_clean$previous_weeks_volume))
data_clean$next_weeks_open <- as.numeric(gsub("[^0-9.]", "", data_clean$next_weeks_open))
data_clean$next_weeks_close <- as.numeric(gsub("[^0-9.]", "", data_clean$next_weeks_close))

# # Convert 'percent_change_next_weeks_price' to numeric by removing '%' and dividing by 100
# data_clean$percent_change_next_weeks_price <- as.numeric(gsub("%", "", data_clean$percent_change_next_weeks_price)) / 100

# Convert 'days_to_next_dividend' to numeric
data_clean$days_to_next_dividend <- as.numeric(gsub("[^0-9.]", "", data_clean$days_to_next_dividend))

str(data_clean)

```

```{r}

# Step 1: Inspect the unique stock symbols
unique_stocks <- unique(data$stock) # Adjust column name if different

# Step 2: Randomly select 5 unique stocks
set.seed(123) # Set seed for reproducibility
selected_stocks <- sample(unique_stocks, 5)

# Step 3: Create a directory for each stock and save its data
for (stock in selected_stocks) {
  # Filter data for the current stock
  stock_data <- data %>%
    filter(stock == !!stock) %>%
    select(stock,date, open, high, low, close, volume) # Adjust column names to match your dataset

  # Create a dedicated folder for the stock
  stock_folder <- here("output_files", paste0("stock_", stock))
  dir.create(stock_folder, showWarnings = FALSE, recursive = TRUE)

  # Save the stock data as a CSV file within its folder
  file_name <- here(stock_folder, paste0("stock_", stock, ".csv"))
  write.csv(stock_data, file_name, row.names = FALSE)
}
```

```{r}

# Use the here function to construct the file path and import the dataset
BA <- read.csv(here("stock_BA.csv"), header = TRUE, sep = ",")
GE <- read.csv(here("stock_GE.csv"), header = TRUE, sep = ",")
INTC <- read.csv(here("stock_INTC.csv"), header = TRUE, sep = ",")
JNJ <- read.csv(here("stock_JNJ.csv"), header = TRUE, sep = ",")
MCD <- read.csv(here("stock_MCD.csv"), header = TRUE, sep = ",")
```